Generating Hubbard Model Solutions from Anderson Impurity Model Solutions

### 2021-08-01 notes.tex
added general Hamiltonian mapping

\title{Generating Hubbard Model Solutions from Anderson Impurity Model Solutions}
\author{Abhirup Mukherjee, Dr. Siddhartha Lal}
\begin{document}
\maketitle
\section{Introduction}
This is an attempt to obtain solutions of the Hubbard model (groundstate wavefunction and ground state energy) using the solutions of the simpler single-impurity Anderson model (SIAM). The models are defined using the Hamiltonians
\begin{equation}\begin{aligned}
H_\text{hubb} &= -t^H\sum_{\sigma,\left<i,j \right>}\left(c^\dagger_{i\sigma} c_{j\sigma} + \text{h.c.}\right) + U^H\sum_i \hat n_{i \uparrow} \hat n_{i \downarrow} - \mu^H \sum_{i\sigma}\hat n_{i\sigma}\\
	H_\text{siam} &= -t^A\sum_{\sigma,\left<i,j \right>\atop{i \neq d \neq j}}\left(c^\dagger_{i\sigma} c_{j\sigma} + \text{h.c.}\right) + \epsilon_d^A \sum_\sigma\hat n_{d\sigma} + U^A \hat n_{d \uparrow} \hat n_{d \downarrow} - \mu^A \sum_{i\neq d,\sigma}\hat n_{i\sigma}
\end{aligned}\end{equation}
Broadly speaking, the method involves first solving the SIAM using a unitary renormalisation group approach, to get the ground state wavefunction and energy eigenvalue, and then combining these wavefunctions in a symmetrized fashion to get the wavefunction for the Hubbard model lattice. It is quite similar to dynamical mean-field theory (DMFT) in essence, but differs in practice. The similar part is the involvement of an impurity-solver. The difference, however, lies in the following points:
\begin{itemize}
	\item While DMFT primarily works with Green's functions and self-energies, this method involves Hamiltonians and wavefunctions.
	\item The impurity-solver in DMFT provides an impurity Green's function (which is then equated with the local Green's function of the bath), while the impurity-solver in this method actually provides a wavefunction.
	\item The final step of DMFT is the self-consistency equation, where the impurity and bath-local quantities are set equal. This ensures all sites, along with the impurity site, have the same self-energy, something which is required on grounds of  translational invariance. The present method, however, brings about the translational invariance in a different way. It symmetrizes the wavefunctions and Hamiltonians itself, such that all quantites then derived from the Hamiltonian or wavefunction are then guaranteed to have the symmetry.
\end{itemize}
The meaning of each of these statements will become clearer when we describe the method in more detail.

@@

The wavefunctions for the \(N=2\) sector can also be connected through these transformations. Since both the Hamiltonians are analytically solvable, we can write down their groundstate wavefunctions:
\begin{equation}\begin{aligned}
	\ket{\Psi^H_\text{GS}} &= 4t^H\frac{1}{\sqrt 2}\left(\ket{\uparrow_1, \downarrow_2} - \ket{\downarrow_1, \uparrow_2}\right) + \left[\Delta\left( U^H, t^H \right) - U^H\right]{\sqrt 2}\left(\ket{\uparrow_1\downarrow_1, } - \ket{,\uparrow_2\downarrow_2}\right)\\
	\ket{\Psi^A_\text{GS}} &= 4t^A\frac{1}{\sqrt 2}\left(\ket{\uparrow_d, \downarrow_z} - \ket{\downarrow_d, \uparrow_z}\right) + \left[\Delta\left( \frac{1}{2}U^A, t^A \right) - \frac{1}{2}U^A\right]{\sqrt 2}\left(\ket{\uparrow_d\downarrow_d, } - \ket{,\uparrow_z\downarrow_z}\right)\\
	E^H_\text{GS} &=  -\frac{1}{2}\Delta\left( U^H, t^H \right),\quad E^A_\text{GS} =  -\frac{1}{2}\Delta\left( \frac{1}{2}U^A, t^A \right), \quad\Delta(x,y) \equiv \sqrt{x^2 + y^2}\\
\end{aligned}\end{equation}
From the forms of the wavefunctions and eigenenergies, we can immediately write down
\begin{equation}\begin{aligned}
	\ket{\Psi^H_\text{GS}} &= \frac{1}{2}\left[\ket{\Psi^A_\text{GS}}\left(t^A, U^A \to t^H, 2U^H, d \to 1, z \to 2\right) + \ket{\Psi^A_\text{GS}}\left(t^A, U^A \to t^H, 2U^H, d \to 2, z \to 1\right)\right]\\
	E^H_\text{GS} &= E^A_\text{GS}\left(t^A, U^A \to t^H, 2U^H\right)
\end{aligned}\end{equation}
This shows that the rules laid out before work for the Hamiltonians, as well as the wavefunctions and energy eigenvalues of the \(N=2,0,4\) sector. These sectors specifically work because it is only in these sectors can we ensure that \(n_d = n_z\), which is required for the Hubbard Hamiltonian because \(n_1 = n_2\). In the other sectors (\(N=1,3\)), the impurity site and the zero-mode sites have to be singly-occupied in some part, and since the impurity site incurs a single-occupation cost of \(-\frac{U^H}{2}\) which is not borne by the zero-mode site, there is an intrinsic dissimilarity between the two sites of the Anderson molecule in this regime. This dissimilarity does not exist for the Hubbard model, so we cannot hope to connect the two models in this regime.

>>>>

\title{Generating Hubbard Model Solutions from Anderson Impurity Model Solutions}
\author{}%Abhirup Mukherjee, Dr. Siddhartha Lal}
\begin{document}
\maketitle
\section{Introduction}
This is an attempt to obtain solutions of the Hubbard model (groundstate wavefunction and ground state energy) using the solutions of the simpler single-impurity Anderson model (SIAM). The models are defined using the Hamiltonians
\begin{equation}\begin{aligned}
H_\text{hubb} &= -t^H\sum_{\sigma,\left<i,j \right>}\left(c^\dagger_{i\sigma} c_{j\sigma} + \text{h.c.}\right) + U^H\sum_i \hat n_{i \uparrow} \hat n_{i \downarrow} - \mu^H \sum_{i\sigma}\hat n_{i\sigma}\\
	H_\text{siam} &= -t^A\sum_{\sigma,\left<i,j \right>\atop{i \neq d \neq j}}\left(c^\dagger_{i\sigma} c_{j\sigma} + \text{h.c.}\right) + \epsilon_d^A \sum_\sigma\hat n_{d\sigma} + U^A \hat n_{d \uparrow} \hat n_{d \downarrow} - \mu^A \sum_{i\neq d,\sigma}\hat n_{i\sigma}
\end{aligned}\end{equation}
Broadly speaking, the method involves first solving the SIAM using a unitary renormalisation group approach, to get the ground state wavefunction and energy eigenvalue, and then combining these wavefunctions in a symmetrized fashion to get the wavefunction for the Hubbard model lattice. It is quite similar to dynamical mean-field theory (DMFT) in essence, but differs in practice. The similar part is the involvement of an impurity-solver. The difference, however, lies in the following points:
\begin{itemize}
	\item While DMFT primarily works with Green's functions and self-energies, this method involves Hamiltonians and wavefunctions.
	\item The impurity-solver in DMFT provides an impurity Green's function (which is then equated with the local Green's function of the bath), while the impurity-solver in this method actually provides a wavefunction.
	\item The final step of DMFT is the self-consistency equation, where the impurity and bath-local quantities are set equal. This ensures all sites, along with the impurity site, have the same self-energy, something which is required on grounds of  translational invariance. The present method, however, brings about the translational invariance in a different way. It symmetrizes the wavefunctions and Hamiltonians itself, such that all quantites then derived from the Hamiltonian or wavefunction are then guaranteed to have the symmetry.
\end{itemize}
The meaning of each of these statements will become clearer when we describe the method in more detail.

@@

The wavefunctions for the \(N=2\) sector can also be connected through these transformations. Since both the Hamiltonians are analytically solvable, we can write down their groundstate wavefunctions \cite{pavarini}:
\begin{equation}\begin{aligned}
	\ket{\Psi^H_\text{GS}} &= 4t^H\frac{1}{\sqrt 2}\left(\ket{\uparrow_1, \downarrow_2} - \ket{\downarrow_1, \uparrow_2}\right) + \left[\Delta\left( U^H, t^H \right) - U^H\right]{\sqrt 2}\left(\ket{\uparrow_1\downarrow_1, } - \ket{,\uparrow_2\downarrow_2}\right)\\
	\ket{\Psi^A_\text{GS}} &= 4t^A\frac{1}{\sqrt 2}\left(\ket{\uparrow_d, \downarrow_z} - \ket{\downarrow_d, \uparrow_z}\right) + \left[\Delta\left( \frac{1}{2}U^A, t^A \right) - \frac{1}{2}U^A\right]{\sqrt 2}\left(\ket{\uparrow_d\downarrow_d, } - \ket{,\uparrow_z\downarrow_z}\right)\\
	E^H_\text{GS} &=  -\frac{1}{2}\Delta\left( U^H, t^H \right),\quad E^A_\text{GS} =  -\frac{1}{2}\Delta\left( \frac{1}{2}U^A, t^A \right), \quad\Delta(x,y) \equiv \sqrt{x^2 + y^2}\\
\end{aligned}\end{equation}
From the forms of the wavefunctions and eigenenergies, we can immediately write down
\begin{equation}\begin{aligned}
	\ket{\Psi^H_\text{GS}} &= \frac{1}{2}\left[\ket{\Psi^A_\text{GS}}\left(t^A, U^A \to t^H, 2U^H, d \to 1, z \to 2\right) + \ket{\Psi^A_\text{GS}}\left(t^A, U^A \to t^H, 2U^H, d \to 2, z \to 1\right)\right]\\
	E^H_\text{GS} &= E^A_\text{GS}\left(t^A, U^A \to t^H, 2U^H\right)
\end{aligned}\end{equation}
This shows that the rules laid out before work for the Hamiltonians, as well as the wavefunctions and energy eigenvalues of the \(N=2,0,4\) sector. These sectors specifically work because it is only in these sectors can we ensure that \(n_d = n_z\), which is required for the Hubbard Hamiltonian because \(n_1 = n_2\). In the other sectors (\(N=1,3\)), the impurity site and the zero-mode sites have to be singly-occupied in some part, and since the impurity site incurs a single-occupation cost of \(-\frac{U^H}{2}\) which is not borne by the zero-mode site, there is an intrinsic dissimilarity between the two sites of the Anderson molecule in this regime. This dissimilarity does not exist for the Hubbard model, so we cannot hope to connect the two models in this regime.

\section{Creating General \(N-\)site Hubbard Hamiltonian from Anderson molecule}
We will follow the strategy outlined in the previous section. For concreteness, we will consider a lattice of \(N\) lattice sites and \(w\) nearest neighbours for each site. Note that a uniform number of nearest neighbours means that there is perfect translational invariance on the lattice, which means there cannot be any edge sites. This is achieved by applying periodic boundary conditions on the edges of the lattice. A square 2d lattice is thus placed on a torus.

For each site \(i\) out of the \(N\) sites (having nearest neighbours \(\{i_j, j\in\left[1,w\right]\}\)), we will create \(w\) local Hamiltonians \(\{H^H_{i,i_j}, i\in \left[1,N\right], j \in \left[1,w\right]  \}\) out of the Anderson molecule Hamiltonian, by setting the impurity index \(d\) to \(i\) and the zero-mode index \(z\) to \(i_j\). We will also need to suitably transform the Hamiltonian parameters \(U^A, t^A\), but we will figure those out as we go along. All \(N\) sites will then produce \(Nw\) local Hamiltonians in total, and the general Hubbard Hamiltonian should be the average of these local Hamiltonians.

The local Hamiltonians look like
\begin{equation}\begin{aligned}
	H^H_{i,i_j} = -t^A\sum_{\sigma}\left(c^\dagger_{i\sigma}c_{i_j\sigma} + \text{h.c.}\right) + U^A\hat \tau_{i \uparrow}\hat \tau_{i \downarrow}
\end{aligned}\end{equation}
We havent transformed the Anderson parameters \(U^A, t^A\) to the Hubbard parameters yet. The sum of all local Hamiltonians for a particular site \(i\) then gives the total Hamiltonian for that site:
\begin{equation}\begin{aligned}
	H^H_{i} = \sum_{j\in\left[1,w\right]}H^H_{i,i_j} = \sum_{\sigma\atop{j \in \text{N.N. of }i}}-t^A\left(c^\dagger_{i\sigma}c_{j\sigma} + \text{h.c.}\right) + w U^A\hat \tau_{i \uparrow}\hat \tau_{i \downarrow}
\end{aligned}\end{equation}
If we now sum over all the lattice sites and take the average of all the Hamiltonians (\(Nw\) in number), we get
\begin{equation}\begin{aligned}
	H^H = \frac{1}{Nw}\sum_{i=1}^N H^H_{i} = -\frac{2}{Nw}t^A\sum_{\sigma\atop{\left<ij \right>}}\left(c^\dagger_{i\sigma}c_{j\sigma} + \text{h.c.}\right) + \frac{1}{N}U^A\hat \tau_{i \uparrow}\hat \tau_{i \downarrow}
\end{aligned}\end{equation}
The factor of 2 in front of \(t^A\) appears because each nearest neighbour pair \(\left<i,j \right>\) is counted twice, once when \(d\to i\) and again when \(d \to j\). Comparing with the actual Hubbard model, we can guess the following transformations for the model paramters:
\begin{equation}\begin{aligned}
	\frac{2}{Nw}t^A \to t^H, && \frac{1}{N}U^A \to U^H
\end{aligned}\end{equation}
For the dimer versions, we had \(N=2, w=1\) and if we substitute these values into the transformations we recover the transformations we had discovered that worked for the dimer.
